Arrows for Parallel Computations

### 2017-12-13 paper/content/arrows.tex
fixes in arrows.tex. reasoning why we chose Arrows over Profunctors

\begin{figure}[h]
	\centering
	\begin{tabular}{cc}
		% \subcaptionbox
{\label{t1}}{\includegraphics[width = 1.5in]{images/first}} &
		% \subcaptionbox
{\label{fig:secondImg}}{\includegraphics[width = 1.5in]{images/second}} \\
|first| & |second| \\
\midrule
		% \subcaptionbox
{}{\includegraphics[width = 1.5in]{images/starstarstar}} &
		% \subcaptionbox
{}{\includegraphics[width = 1.5in]{images/andandand}}\\
|(***)|\label{fig:***Img} & |(&&&)| \label{fig:&&&Img} \\
	\end{tabular}
	\caption{Visual depiction of syntactic sugar for Arrows.}
	\label{fig:syntacticSugarArrows}
\end{figure}
Hughes also defined some syntactic sugar (Fig.~\ref{fig:syntacticSugarArrows}): |second|, |***| and |&&&|. 
|second| is the mirrored version of |first| (Appendix~\ref{utilfns}).
|***| combines |first| and |second| to handle two inputs in one arrow, and is defined as follows:
% \begin{figure}[h]
\begin{code}
(***) :: Arrow arr => arr a b -> arr c d -> arr (a, c) (b, d)
f *** g = first f >>> second g
\end{code}
% \caption{The (***) combinator}
% \label{fig:***}
% \end{figure}
The |&&&| combinator, which constructs an Arrow that outputs two different values like |***|, but takes only one input, is:
% \begin{figure}[h]
\begin{code}
(&&&) :: Arrow arr => arr a b -> arr a c -> arr a (b, c)
f &&& g = arr (\a -> (a, a)) >>> (f *** g)
\end{code}
% \caption{The (\&\&\&) combinator}
% \label{fig:&&&}
% \end{figure}
A~first short example given by Hughes on how to use arrows is addition with arrows:
%|add| over arrows, which can be seen in Fig.~\ref{fig:addArrows}.
% \begin{figure}[h]
\begin{code}
add :: Arrow arr => arr a Int -> arr a Int -> arr a Int
add f g = (f &&& g) >>> arr (\(u, v) -> u + v)
\end{code}
% \caption{Add over arrows}
% \label{fig:addArrows}
% \end{figure}
% The benefit of using the |Arrow| typeclass is that any type which is shown to be an arrow can be used in conjunction with this newly created |add| combinator. Even though this example is quite simple, the power of the arrow interface immediately is clear: If a type is an arrow, it can automatically used together with every library that works on arrows. Compared to simple Monads, this enables us to write code that is more extensible, without touching the internals of the specific arrows.
% \\\\
% \textit{Note: In the definitions Hughes gave in his paper, the notation |a b c| for an arrow from |b| to |c| is used. We use the equivalent definition |arr a b| for an arrow from |a| to |b| instead, to make it easier to find the arrow type in type signatures.}
%

>>>>

\begin{figure}[h]
	\centering
	\begin{tabular}{cc}
		% \subcaptionbox
{\label{t1}}{\includegraphics[width = 1.5in]{images/first}} &
		% \subcaptionbox
{\label{fig:secondImg}}{\includegraphics[width = 1.5in]{images/second}} \\
|first| & |second| \\
\midrule
		% \subcaptionbox
{}{\includegraphics[width = 1.5in]{images/starstarstar}} &
		% \subcaptionbox
{}{\includegraphics[width = 1.5in]{images/andandand}}\\
|(***)|\label{fig:***Img} & |(&&&)| \label{fig:&&&Img} \\
	\end{tabular}
	\caption{Visual depiction of syntactic sugar for Arrows.}
	\label{fig:syntacticSugarArrows}
\end{figure}
Hughes also defined some syntactic sugar (Fig.~\ref{fig:syntacticSugarArrows}): |second|, |***| and |&&&|. 
|second| is the mirrored version of |first| (Appendix~\ref{utilfns}).
|***| combines |first| and |second| to handle two inputs in one arrow, and is defined as follows:
% \begin{figure}[h]
\begin{code}
(***) :: Arrow arr => arr a b -> arr c d -> arr (a, c) (b, d)
f *** g = first f >>> second g
\end{code}
% \caption{The (***) combinator}
% \label{fig:***}
% \end{figure}
The |&&&| combinator, which constructs an Arrow that outputs two different values like |***|, but takes only one input, is:
% \begin{figure}[h]
\begin{code}
(&&&) :: Arrow arr => arr a b -> arr a c -> arr a (b, c)
f &&& g = arr (\a -> (a, a)) >>> (f *** g)
\end{code}
% \caption{The (\&\&\&) combinator}
% \label{fig:&&&}
% \end{figure}
A~first short example given by Hughes on how to use arrows is addition with arrows:
%|add| over arrows, which can be seen in Fig.~\ref{fig:addArrows}.
% \begin{figure}[h]
\begin{code}
add :: Arrow arr => arr a Int -> arr a Int -> arr a Int
add f g = (f &&& g) >>> arr (\(u, v) -> u + v)
\end{code}

As we can rewrite the monadic bind operation |(>>=)| with only the Kleisli type into |m a -> Kleisli m a b -> m b|, but not with a general Arrow |arr a b|, we can intuitively get an idea of why Arrows must be a generalisation of Monads. While this also means that a general Arrow can not express everything a Monad can, the Parser example shows that this trade-off is worth it in some cases.

In this paper we will show that we can express parallel computations with Arrows and do not require Monads to do so. We also do not restrict the compatible Arrows to ones which have |ArrowApply| instances but only require the |ArrowChoice| and |ArrowLoop|. Because of this, we have a truly more general interface as compared to a monadic one.

Also note that, while we could have based our DSL on Profunctors as well, we chose Arrows. for this paper since they they are a more direct translation of pure functions than Profunctors. Note that Profunctors can easily be adapted to our interface as shown in Appendix~\ref{app:profunctorArrows}. However, they are a promising candidate for future improvement of our DSL.

% \caption{Add over arrows}
% \label{fig:addArrows}
% \end{figure}
% The benefit of using the |Arrow| typeclass is that any type which is shown to be an arrow can be used in conjunction with this newly created |add| combinator. Even though this example is quite simple, the power of the arrow interface immediately is clear: If a type is an arrow, it can automatically used together with every library that works on arrows. Compared to simple Monads, this enables us to write code that is more extensible, without touching the internals of the specific arrows.
% \\\\
% \textit{Note: In the definitions Hughes gave in his paper, the notation |a b c| for an arrow from |b| to |c| is used. We use the equivalent definition |arr a b| for an arrow from |a| to |b| instead, to make it easier to find the arrow type in type signatures.}
%

### 2017-12-13 paper/content/utilityFunctions.tex
fixes in arrows.tex. reasoning why we chose Arrows over Profunctors

>>>>

\section{Profunctor Arrows}
\label{app:profunctorArrows}

Following is a theoretical way of using Profunctors together with our DSL.
\footnote{for additional information on the typeclasses, see: \url{https://hackage.haskell.org/package/profunctors-5.2.1/docs/Data-Profunctor.html}}

\begin{code}
instance (Category p, Strong p) => Arrow p where
  arr f = dimap id f id
  first = first'

instance (Category p, Strong p, Costrong p) => ArrowLoop p where
  loop = loop'

instance (Category p, Strong p, Choice p) => ArrowChoice p where
  left = left'
\end{code}
